---
title: "Library Acquisitions Over Time"
author: "Marie Jordan"
output: pdf_document
---

```{r include=FALSE}
library(tidyverse)
library(ggplot2)
library(lubridate)
library(scales)
```
# Data Source

All datasets were downloaded from the IMLS website on 7/30/2019.

https://www.imls.gov/research-evaluation/data-collection/public-libraries-survey

Relevant context from IMLS:

Coverage: The data are collected from approximately 9,000 public libraries with approximately 17,000 individual public library outlets (main libraries, branches, and bookmobiles) in the 50 states, the District of Columbia, and outlying territories.

Methods: At the state level, PLS is administered by Data Coordinators, appointed by the chief officer of the state library agency from each state or outlying area. State Data Coordinators collect the requested data from local public libraries and report these data to us via a web-based reporting system.

Further documentation is available with each dataset.

```{r include=FALSE}
fy2017 <- read_csv('PLS_FY2017_Data_Files_CSV/PLS_FY17_AE_pud17i.csv')
fy2016 <- read_csv('pls_fy2016_data_files_csv/PLS_FY2016_AE_pupld16a_updated.csv')
```

```{r include=FALSE}
fy2017_sm <- subset(fy2017,select=c('PRMATEXP','ELMATEXP','OTHMATEX','TOTEXPCO'))
fy2016_sm <- subset(fy2016, select=c('PRMATEXP','ELMATEXP','OTHMATEX','TOTEXPCO'))
```

```{r include=FALSE}
fy2017_sm$YEAR <- '2017'
fy2016_sm$YEAR <- '2016'
```

```{r include=FALSE}
fy_comp <- rbind(fy2017_sm,fy2016_sm)
```



```{r include=FALSE}
fy_comp_sum <- fy_comp %>% 
        group_by(YEAR) %>% 
        summarise(total_print = sum(PRMATEXP),total_electronic = sum(ELMATEXP),total_other=sum(OTHMATEX))
```

```{r include=FALSE}
fy_comp_sum$total_print <- fy_comp_sum$total_print / 100000000
fy_comp_sum$total_electronic <- fy_comp_sum$total_electronic / 100000000
fy_comp_sum$total_other <- fy_comp_sum$total_other / 100000000
```

```{r include=FALSE}
fy_comp_sum$YEAR <- as.integer(fy_comp_sum$YEAR)

```
# Total Expenditures By Year
```{r echo=FALSE}
print(fy_comp_sum)
```
Note that numbers are in millions, e.g. in 2016, 7.5 million dollars went to print
expenditures.

```{r echo=FALSE}
ggplot(fy_comp_sum,aes(YEAR,total_print,color='Print Total Expenditure'))+
  geom_line()+
  geom_line(data=fy_comp_sum, aes(YEAR,total_electronic,color='Electronic Total Expenditure'))+
  geom_line(data=fy_comp_sum, aes(YEAR,total_other,color='Other Total Expenditure')) +
  labs(title="Collection Expenditures")+
  labs(x = "Year", y = "Sum of Expenditure (in millions)")+
  scale_color_manual(values = c('Print Total Expenditure' = 'black',
    'Electronic Total Expenditure' = '#c89fe0', 'Other Total Expenditure'='#308025'))
```




